---
title: "Data Cleanup"
author: "Vuthy (From Sekhar's Code)"
date: "Sunday, March 29, 2015"
output: html_document
---
```{r echo=FALSE, results='hide', message=FALSE}
require("RCurl")
url <- "https://raw.githubusercontent.com/srajeev1/607Project3/master/Awards_DB/Detailed_Analysis_Read_if_interested/academy_awards.csv"
myData <- getURL(url, ssl.verifypeer=FALSE)
academy_awards_df <- read.csv(textConnection(myData), header=TRUE,stringsAsFactors=F)
str(academy_awards_df)
````
#### Data Issues:
1. Additional Variables 
2. Unwated characters in data
3. Movie Names and nominees were interchanged for the following categories

* Actor -- Leading Role
* Actress -- Leading Role 
* Actor -- Supporting Role
* Actress -- Supporting Role

4. Obsolete Categories

````{r}
#Delete Additional Variables
academy_awards_df$X <-NULL
academy_awards_df$X.1 <-NULL
academy_awards_df$X.2 <-NULL
academy_awards_df$X.3 <-NULL
academy_awards_df$X.4 <-NULL
academy_awards_df$X.5 <-NULL

````

```{r message=FALSE}
library(dplyr)
library(tidyr)
```


Remove the text between () in Year column
```{r}
#Seperate the year column into two sections
academy_awards_df <- academy_awards_df %>%
  separate(Year,into=c("Year","Junk"), sep="\\(")

#Delete the section (obsolete) sections
academy_awards_df$Junk <- NULL

#Seperate the year column into two sections
academy_awards_df <- academy_awards_df %>%
  separate(Year,into=c("Year","Junk"), sep="\\ ")

#Delete the section (obsolete) sections
academy_awards_df$Junk <- NULL
```


Extract the correct year part from Years, which are separated by a "/".
Older records contained data in the following format: 1927/28 (1st)
```{r}
#If the length of 'Years' equal 7, take only the last two characters and add '19' to the front of it.
academy_awards_df$Year <- ifelse(nchar(academy_awards_df$Year) == 7,paste("19",substr(academy_awards_df$Year,6,7),sep=""),academy_awards_df$Year)

```

```{r}
#Removing the text between "{}" from Additional.Info variable
academy_awards_df$Additional.Info <- as.vector(substring(academy_awards_df$Additional.Info,1,
        ifelse(regexpr("\\{",academy_awards_df$Additional.Info) == -1, nchar(academy_awards_df$Additional.Info),
               (regexpr("\\{",academy_awards_df$Additional.Info) - 1))
)
)
```

```{r}
#Finding the movie name from Nominee and Additional.Info variables, since for some award categories the information was interchanged
academy_awards_df$Movie <- ifelse((
  academy_awards_df$Category == "Actor -- Leading Role" | 
  academy_awards_df$Category == "Actor -- Supporting Role" | 
  academy_awards_df$Category == "Actress -- Leading Role" | 
  academy_awards_df$Category == "Actress -- Supporting Role"), 
                                  academy_awards_df$Additional.Info, academy_awards_df$Nominee
)

#Now that we have obtained the Movie name, delete the unnecessary variables
academy_awards_df$Additional.Info <- NULL
academy_awards_df$Nominee <- NULL
```


```{r}
#Changing the Won. variable to Won, and replacing the YES with 1 and NO with 0
academy_awards_df$Won <- ifelse(academy_awards_df$Won. == "NO", 0, 1)
academy_awards_df$Won. <- NULL
```


```{r}
#Delete the records belonging to the following award categories, since these awards are not associated to any movie or archaic:

#Honorary Award
#Irving G. Thalberg Memorial Award
#Scientific and Technical (Scientific and Engineering Award)
#Scientific and Technical (Technical Achievement Award)
#Scientific and Technical (Bonner Medal)
#Jean Hersholt Humanitarian Award
#Scientific and Technical (Special Awards)
#Scientific and Technical (Academy Award of Merit)
#Scientific and Technical (Gordon E. Sawyer Award)
#Acting (other)

#Engineering Effects (archaic category)
#Unique and Artistic Picture (archaic category)
#Special Effects (archaic category)
#Dance Direction (archaic category)
#Assistant Director (archaic category)
#Foreign Language Film

academy_awards_df <- academy_awards_df[ -which((
  academy_awards_df$Category == "Irving G. Thalberg Memorial Award" | 
  academy_awards_df$Category == "Scientific and Technical (Scientific and Engineering Award)" | 
  academy_awards_df$Category == "Scientific and Technical (Technical Achievement Award)" |    
  academy_awards_df$Category == "Scientific and Technical (Bonner Medal)" |
  academy_awards_df$Category == "Honorary Award" |
  academy_awards_df$Category == "Jean Hersholt Humanitarian Award" |
  academy_awards_df$Category == "Scientific and Technical (Special Awards)" |
  academy_awards_df$Category == "Scientific and Technical (Academy Award of Merit)" |
  academy_awards_df$Category == "Scientific and Technical (Gordon E. Sawyer Award)" |
  academy_awards_df$Category == "Special Achievement Award" |
  academy_awards_df$Category == "Acting (other)" |
  academy_awards_df$Category == "Engineering Effects (archaic category)" |
  academy_awards_df$Category == "Unique and Artistic Picture (archaic category)" |
  academy_awards_df$Category == "Special Effects (archaic category)" |
  academy_awards_df$Category == "Dance Direction (archaic category)" |
  academy_awards_df$Category == "Assistant Director (archaic category)" |
  academy_awards_df$Category == "Foreign Language Film"                                          
)),]

```

```{r}
#Create Trim functions
trim.leading <- function (x)  sub("^\\s+", "", x)
trim.trailing <- function (x) sub("\\s+$", "", x)
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```

Generate CSV files for loading into the RDBMS Tables:

Creating the file for Category table:
```{r}
academy_awards_df$Category <- trim(academy_awards_df$Category)
academy_awards_df$Category <- toupper(academy_awards_df$Category)

write.table(data.frame(unique(academy_awards_df$Category)),file="Category.csv",row.names=FALSE,col.names=FALSE,sep=",",quote=FALSE)
```

Creating the file for Movie table:
```{r}
academy_awards_df$Movie <- trim(academy_awards_df$Movie)
academy_awards_df$Movie <- toupper(academy_awards_df$Movie)

write.table(unique(data.frame(academy_awards_df$Movie,academy_awards_df$Year)) ,file="Movie.csv",row.names=FALSE,col.names=FALSE,sep=",",quote=TRUE)

write.table(academy_awards_df ,file="Academy_Awards_Temp.csv",row.names=FALSE,col.names=FALSE,sep=",",quote=TRUE)
````

After the data files are generated, it was loaded into a database and a select query was issued to generate the data the final data file.