---
title: "Data Visualization"
author: "Honey, Vuthy, James, Sonya, Mohamed"
date: "March 24, 2015"
output: html_document
---

Data Visualization for the IS607 Project No. 2 | CUNY MSDA 

Load data
```{r}
load("Awards.Data.RData")

#View(data.awards)
```

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)
#library(grid)
library(gridExtra)
```

## Table 1. Number of times a Best Picture winner was nominated or won another category. 
```{r, warning=FALSE, message=FALSE}
# Create BP, a dataframe showing only Best Picture Winners
BP <- filter(data.awards, category_name == "BEST PICTURE", won == 1)

# Create BP.AlsoNom, a dataframe showing what other categories the Best Picture Winners were NOMINATED for
BP.AlsoNom <- left_join(BP, data.awards, by = "movie_name") 
BP.AlsoNom <- filter(BP.AlsoNom, category_name.y != "BEST PICTURE") 
BP.AlsoNom <- select(BP.AlsoNom, category_name.y) 
BP.AlsoNom <- count(BP.AlsoNom, category_name.y)
BP.AlsoNom <- arrange(BP.AlsoNom, desc(n))

# Create BP.AlsoWon, a dataframe showing what other categories the Best Picture winners WON
data.awards.won <- filter(data.awards, won == 1)

BP.AlsoWon <- left_join(BP, data.awards.won, by = "movie_name")
BP.AlsoWon <- filter(BP.AlsoWon, category_name.y != "BEST PICTURE") 
BP.AlsoWon <- select(BP.AlsoWon, category_name.y) 
BP.AlsoWon <- count(BP.AlsoWon, category_name.y)
BP.AlsoWon <- arrange(BP.AlsoWon, desc(n))

kable(cbind(BP.AlsoNom, BP.AlsoWon), 
      col.names = c("Category Nominations", "Count", "Category Winners", "Count"))
```

##Figure 1. Barchart showing the number of times a Best Picture winner was nominated or won another category. 
```{r}
nom <- ggplot(data = BP.AlsoNom, aes(x = reorder(category_name.y, n), y=n)) + geom_bar(stat="identity") + coord_flip() + labs(x = "Category", y = "Nominations")

win <- ggplot(data = BP.AlsoWon, aes(x = reorder(category_name.y, n), y=n)) + geom_bar(stat="identity") + coord_flip() + labs(x = "Category", y = "Wins") + ylim(0, 80)

grid.arrange(nom, win, ncol=1)
```

## Correlation Matrices

Use Sekhar's method to create award and category matrices
```{r, echo=TRUE}
library(dplyr)
library(tidyr)

awards_df <- read.csv("Awards_File.csv",stringsAsFactors=F)
awards_modified <- awards_df[,c(1,3,4,6)]
head(awards_modified)
awards_modified <- awards_modified %>%
  group_by(movie_id,year,category_id) %>%
  summarise(won=max(won))
awards_re_modified <- spread((awards_modified),category_id,won)
head(awards_re_modified)
award_categories <- unique(data.frame(category_id=awards_df$category_id,category_name=awards_df$category_name))
rownames(award_categories) <- NULL
names(awards_re_modified) <- c("movie_id", "year",  "c1",  "c2",  "c3",  "c4",  "c5", "c6",  "c7",  "c8",  "c9", "c10", "c11", "c12", "c13", "c14", "c15","c16", "c17", "c18", "c19", "c20", "c21", "c22", "c23")
```

Modify awards matrix to set NA = 0
```{r, echo = TRUE}
####
names(awards_re_modified) <- c("movie_id", "year","ACTOR -- LEADING ROLE","ACTOR -- SUPPORTING ROLE","ACTRESS
LEADING ROLE","ACTRESS SUPPORTING ROLE","ANIMATED FEATURE FILM","ART DIRECTION","CINEMATOGRAPHY","COSTUME
DESIGN","DIRECTING","DOCUMENTARY FEATURE","DOCUMENTARY SHORT SUBJECT","FILM EDITING","MAKEUP","MUSIC
SCORING","MUSIC SONG","BEST PICTURE","SHORT FILM ANIMATED","SHORT FILM LIVE ACTION","SOUND","SOUND
EDITING","VISUAL EFFECTS","WRITING","DOCUMENTARY OTHER")
awards_mod <- awards_re_modified #create new var to preserve original
awards_mod[is.na(awards_mod)] <- 0 #set NA = 0
```

## Figure 2. Correlation matrix, display circles

```{r, echo = TRUE}

library(corrplot)
W <- cor(awards_mod[3:25]) #correlate c1:c23
corrplot(W, method = "circle", tl.cex = 0.5)
```

## Figure 3. Correlation matrix, display ellipses
```{r, echo=TRUE}
corrplot(W, method = "ellipse", tl.cex = 0.5)
```

## Figure 4. Correlation matrix, display squares
```{r, echo=TRUE}
corrplot(W, method = "square", tl.cex = 0.5)
```

## Figure 5. Correlation matrix, using ggplot2
```{r, echo=TRUE}
library(ggplot2)
library(reshape2)
cm <- qplot(x=Var1, y=Var2, data=melt(cor(awards_mod[3:25])), fill=value, geom="tile")
cm + theme(axis.text=element_text(size=8))
```

## Table 2. Correlation Matrix, numerical values
```{r, echo=TRUE}
x <- awards_mod[3:25]
y <- awards_mod[3:25]
cor(x, y)
```

***An analysis of the various correlation matrix visualizations and table of numerical values shows the
primary determining factor is DIRECTING, followed by WRITING and then FILM EDITING.***

Note: There was an issue with NA values; we could not get them to be ignored, so instead had to set them
equal to 0, which may have resulted in some skewing of the results.

```{r}
require(ggplot2)
library(ggplot2)
library(plot3D)
library(scatterplot3d)

```

 I took the summary table that Honey, Vuthy, James, and Sonya had created and added few graphics as well

 Please note that i only took the summary with the below following columns:
Category.Nominations,
Count,
Category.Winners,
Count

 Also, I loaded and shared the table as csv using dropconvas website as per below 

```{r}

data1<- read.table(file= "http://dropcanvas.com/dcm0h/1", header = TRUE, sep=",")
str(data1)
head(data1)


```


# Distribution Nominations By Category

```{r, echo=FALSE}

p<- qplot(Category.Nominations, Count,  data= data1, color =Category.Nominations,  size=I(5),
           main = " Distribution Nominations By Category",
           xlab = " Category Nominations ", 
           ylab = " Nomination Count " )   

p + theme_bw()
p


```

# Distribution of Winning By Category

```{r, echo=FALSE}


p<- qplot(Category.Winners, Count.1,  data= data1, color =Category.Winners,  size=I(5),
          main = " Distribution of Winning By Category ",
          xlab = " Winning Category ", 
          ylab = " Winning Count " )   

p + theme_bw()
p

```

# Distribution of Wins over Winning by Nominations Category

```{r}
attach(data1) 
scatterplot3d(Category.Nominations, Category.Winners, Count.1, pch=16, highlight.3d=TRUE,
               type="h", main="Distribution of Wins over Winning by Nominations Category",
               xlab = "Category.Nominations", ylab = "Winning Category", zlab = " Winners")

```

