---
title: "Project_3_Report"
author: "Sekhar Mekala"
date: "Monday, March 23, 2015"
output: html_document
---

# Data Massaging

```{r}
library(dplyr)
library(tidyr)

setwd("C:/Users/Sekhar/Documents/Github/Project_3_607_My_Work")
awards_df <- read.csv("Awards_File.csv",stringsAsFactors=F)
awards_modified <- awards_df[,c(1,3,4,6)]
#head(awards_modified)
```

The awards_modified data frame has some movies, which are nominated for the same category more than once, and has also won for one of its nominations (see below example) 
```{r}
awards_modified[awards_modified$movie_id==1288,]
```

To fix the above problem, we have to group by movie_id, year, category_id, and max(won)
such grouping of data will include just the winning nomination, if a film is nominated multiple times in the same category, and one of the nominations wins in that category.

```{r}
awards_modified <- awards_modified %>%
  group_by(movie_id,year,category_id) %>%
  summarise(won=max(won))
```

The above transformation has eliminated all the duplicate rows, where the same movie has been nominated in the same category more than once and one of them wins.
It just includes the winning nomination. If none of the multiple nominations win, then only one of the nominations are included.

The below command confirms that for the movie_id=1536, only the winning nomination is inluded
```{r}
awards_modified[awards_modified$movie_id==1288,]
```

Now the spread() function can be applied to awards_modified data frame.
```{r}
awards_re_modified <- spread((awards_modified),category_id,won)
```
The above command will produce another data frame called *awards_re_modified*, with the following columns:
*movie_id* - Movie IDentifier
*Year* - Year of award
*1 to 23* - Award categories from 1 to 23
Wherever a movie wins an award, the respective category will have 1, wherever a movie is nominated, then the corresponding category will have 0, and wherever the movie  is neither nominated nor wins, we will have NA

```{r}
head(awards_re_modified)
```

To get the category names (since the above dataframe contains the categories as numbers/codes), let us create a separate data frame *award_categories*:

```{r}
award_categories <- unique(data.frame(category_id=awards_df$category_id,category_name=awards_df$category_name))
rownames(award_categories) <- NULL

award_categories
```

Since the variables are not allowed to start with numeric values, let us rename the column names of the data frame awards_re_modified to character variables.
For example, the variable name "1" represents the award category_ID 1. We will change this to c1, to represent category 1.

```{r}
names(awards_re_modified) <- c("movie_id", "year",  "c1",  "c2",  "c3",  "c4",  "c5", "c6",  "c7",  "c8",  "c9", "c10", "c11", "c12",
                               "c13", "c14", "c15","c16", "c17", "c18", "c19", "c20", "c21", "c22", "c23")

```

So finally we have the following data frames to work/use for our analysis:
*award_categories*
*awards_re_modified*



#Main Objectives
The *Best Picture* category_id is *16* and *Film Editing* category_id is *12*. 
We have to identify which categories can help us to predict if a film gets the best picture, and determine if film_editing has the maximum probability.

Our specific objectives are:

* Probability of getting Best Movie, given that the movie is nominated in a paricular award category?
* Probability of getting Best Movie, given that the movie wins in a paricular award category?


#Data Analysis
Let us use conditional probability to determine the following:

**Objective-1** Probability of getting Best Movie, given that the movie is nominated in a paricular award category?

**Objective-1** Probability of getting Best Movie, given that the movie wins in a paricular award category?


## Bayes formula to find conditional probability:
The formula for conditional probability is given below:
$$P(A|B) = P(B|A) . P(A)/P(B) = P(AB) / P(B)$$


In other words .... we will use the following cormulas to satisfy the objectives of our analysis.
For **Objective-1**:

P(Best Movie | It is nominated in x category) = P(that the movie won the Best picture award and also nominated in x category) / P(that a movie is nominated in x category)

For **Objective-2**:

P(Best Movie | (It is nominated in x category and also wins award in x category)) = P(that the movie won the Best picture award and got nominated in x category and also wins award in x category) / P(that a movie is nominated in x category and also wins the award in x category)


Let us create a **temp** datafame, and read all the awards information, dropping the movie_id and year. We will also move the c16 (Best picture) award category as the last variable in **temp** data frame. This will help us to iterate via the columns easily, using for loop construct of R:

```{r}
temp <- data.frame()
temp <- data.frame(awards_re_modified[,3:25])
temp$c0 <- temp$c16
temp$c16 <- NULL
names(temp)[23] <- "c16"

#probability_if_nominated <- data.frame(norow=22,ncol=2)
```

Let us display some sample rows of temp:

```{r}
head(temp)
```

The **R** code to get the conditional probabilities of getting the best picture, given that the movie is nominated in a specific category is given below. 
This code produces a data frame, with the following details:


*Award_Category_ID

*Award_Category_Name

*Probability_Percent


The third variable (Probability_Percent) represents the probability of getting best picture award if the movie is nominated in a specific award category.
##R Code for Objective-1
```{r}
p <- vector(length=22)
category_id <- vector(length=22)
category_name <- vector(length=22)
category_name <- NULL
for(i in 1:22)
{

  x <- temp[,c(i,23)]
  #Get the probability of getting a movie nominated in category i
  #First get the number of movies, and then get the number of movies for which the category_i is 1 or 0
  
  category_nominated_count <- sum(!is.na(x[,1])) #if category_i is 0 or 1, then the movie is nominated
  Exhaustive_movies_count <- nrow(x)
  #Hence the probability of getting a movie nominated is ...
  p_of_getting_nom_in_i <- (category_nominated_count / Exhaustive_movies_count)
  
  
  #Now get the count of movies which won the best movie given that they were nominated in category_i
  p_of_best_pic_and_get_nom_in_i <- (sum(!is.na((x[which(x[,2] == 1),])[1])) / nrow(x))
      
  #Now, the probability of getting best movie given that it is nominated in category_i
   p[i] <- (p_of_best_pic_and_get_nom_in_i / p_of_getting_nom_in_i) 
  category_id[i] <- as.numeric(substr(names(temp)[i],2,3))
  category_name[i] <- as.vector(award_categories[award_categories$category_id==(category_id[i]),2])
  
  
}

prob_of_best_pic_given_cat_nom <- data.frame(Award_Category_ID=category_id,Award_Category_Name=category_name,Probability_Percent=(p*100))
prob_of_best_pic_given_cat_nom 

barplot(prob_of_best_pic_given_cat_nom$Probability_Percent,names.arg=prob_of_best_pic_given_cat_nom$Award_Category_ID, 
        args.legend = prob_of_best_pic_given_cat_nom$Award_Category_Name)
```


The above display of the probabilities conclude that if a film is nominated in "Directing" category, then there is 19% probability of winning the Best Picture award. The "Directing" category is the best predictor for "Best Picture" award. The second best predictor is "Film Editing", which predicts that if a film is nominated for "Film Editing" category, then there is 17.7% probability that it wins "Best Picture" award. 


Let us compute the probabilities of getting "Best Picture" award to a movie, given that the movie wins award in a specific category (*Objective_2*)


The *R* code to get the conditional probabilities of getting the best picture, given that the movie wins in a specific category is given below. 
This code produces a data frame, with the following details:


*Award_Category_ID

*Award_Category_Name

*Probability_Percent

The third variable (Probability_Percent) represents the probability of getting best picture award if the movie is wins in a specific award category.
##R Code for Objective-2

```{r}
for(i in 1:22)
{
  
  x <- temp[,c(i,23)]
  #Get the number of times a movie won in cat_i
  num_of_wins_in_cat_i <- sum(x[which(!is.na(x[,1])),1] == 1)
  
  #Get the number of times a movie got Best movie and also won cat_i
  num_of_best_movies_given_cat_i_win <- sum(x[which(x[,2] == 1),1] == 1,na.rm=T)
  
  #Probability of winning Best pic, given a picture wins in cat_i
  p[i] <- (num_of_best_movies_given_cat_i_win / num_of_wins_in_cat_i)
  
  category_id[i] <- as.numeric(substr(names(temp)[i],2,3))
  category_name[i] <- as.vector(award_categories[award_categories$category_id==(category_id[i]),2])
  
}


prob_of_best_pic_given_a_cat_wins <- data.frame(Award_Category_ID=category_id,Award_Category_Name=category_name,Probability_Percent=(p*100))
prob_of_best_pic_given_a_cat_wins

barplot(prob_of_best_pic_given_a_cat_wins$Probability_Percent,names.arg=prob_of_best_pic_given_a_cat_wins$Award_Category_ID, 
        args.legend = prob_of_best_pic_given_a_cat_wins$Award_Category_Name)

#barplot(prob_of_best_pic_given_cat_nom$Probability_Percent,names.arg=prob_of_best_pic_given_cat_nom$Award_Category_ID, 
#        args.legend = prob_of_best_pic_given_cat_nom$Award_Category_Name)

```

The above display concludes that, if a movie wins the "Directing" award, then there is 71.8% probability that it also wins "Best Picture", and if a film wins "Film Editing" award, then there is 42.8% chance that it can also win "Best Film". Hence "Directing" category predicts the best picture with a maximum probability.

#Final conclusion
The "Driecting" award predicts the "Best Movie" award with maximum probability, and the second best predictor for "Best Film" is the "Film Editing" category
